---
title: "d18o_analysis"
author: "Zack Grzywacz"
date: "June 16, 2021"
output: html_document
---


```{r}
library(lubridate)
library(lattice)
library(rnaturalearth)
library(rnaturalearthdata)
library(rasterVis)
library(seas)
library(s2dverification)
library(reshape2)
library(dplyr)
library(Hmisc)
library(ggplot2)
library(viridis)
library(ggExtra)
library(grid)
library(pheatmap)
library(treeclim)
library(raster)
library(dplR)
library(scPDSI)
```


** Reading in data **

```{r}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
```

```{r}
raw <- read.csv("raw_isotopes.csv") # Raw isotope data
raw$year <- substrRight(raw$sample_ID, 4)
raw$year <- strtoi(raw$year)

prysm <- read.csv("prysm_output.csv") # pysm output

annual_summary <- raw %>% group_by(year) %>% summarise(mean = mean(o18corr, na.rm = TRUE)) # Combined Isotope data

tab <- read.csv("tabulated_18o.csv") # Oxygen data tabulated
rownames(tab) <- tab[,1]
tab[,1] <- NULL

tab2 <- read.csv("tabulated_13c.csv") # carbon data tabulated
rownames(tab2) <- tab2[,1]
tab2[,1] <- NULL

gnip <- read.csv("CG_GNIP.csv")
gnip$Date <- parse_date_time(gnip$Date, "mdy")
gnip$month <- month(gnip$Date)
gnip$year <- year(gnip$Date)
for (i in c(1:288)){
if (gnip$month[i] < 9){
  gnip$seas.year[i] = (gnip$year[i] - 1)
}
else{
  gnip$seas.year[i] = gnip$year[i]
}
}

gnip_outlier <- gnip
gnip_outlier$O18[gnip_outlier$O18 > 1] <- NA

silo <- read.csv("silo_monthly_1958_2019.csv")
```

**Removal of flyers**

Removing flyers:
1. Is it an outlier for voltage? (>5) If yes:
2. Is it an outlier for BOTH O18 and C13? If yes, remove it
To remove:
619 1974
619 1978
609 1982
Maybe: 606 1997

Just removed 1973-1978 for 619. Might undo and just remove 74,78
606 and 607 are a little high in voltage across the board. Just to note 

Look into removing low 605 (1964 or so) due to LOW voltage. Possibly same for 603 1975, 611 1962


```{r}
ggplot(data=tab, aes(x=rownames(tab), y=X611)) +
  geom_line()+
  geom_point()
```


** Ring width detrending then correlations: **

```{r}
master <- read.rwl("MRD6_master.rwl")
```

```{r}
detrends <- apply(master, MARGIN = 2, detrend.series, method="Spline")
detrend_df <- as.data.frame(detrends)
ids <- read.ids(detrend_df, stc=c(4,2,1))
treemeans <- treeMean(detrend_df, ids, na.rm=FALSE)
```

```{r}
ring_means <- treemeans[653:711,] # Ring means for just 1960-2018
```

```{r}
pheatmap(as.matrix(ring_means), scale="none", cluster_rows = FALSE, cluster_cols = FALSE)
```
Change this to see corrs between each tree's isotopes and its rings:
```{r}
cor.test(ring_means[['3']], tab2[['X603']])
```

Corrs are negative. Only some are highly correlated between 18o and ring width.
Carbon corrs are also negative, more are significant


** Comparing & Correlating 18O and 13C series between trees **

Oxygen:
```{r}
pheatmap(as.matrix(tab), scale="none", cluster_rows = FALSE, cluster_cols = FALSE)
```
Carbon:
```{r}
pheatmap(as.matrix(tab2), scale="none", cluster_rows = FALSE, cluster_cols = FALSE)
```

Oxygen correlation matrix:

```{r}
cor(tab, use = "complete.obs")
```
603 - 619, 612, 605
619 - 603, 605, 612
612 - 619, 603

605 - 619, 604, 603 ---------------------Link between groups maybe?

604 - 607, 609, 605
607 - 604, 611
609 - 604
611 - 607

606 - 617**
617 - 606**





Carbon correlation matrix:

```{r}
cor(tab2, use = "complete.obs")
```
603 - 619, 605, 612, 617
619 - 603, 605, 612, 617
605 - 619, 603
612 - 619, 603

617 - 607, 611, 606, 619, 603  -----------617 Links these two groups

607 - 617, 611, 606
606 - 607, 617, 604
611 - 617, 607

604 - 609, 606  -----------------These correlate with each other but negative with most others
609 - 604

Corrs generally higher for 13C than 18O could be because of trend

** Comparing to soil depth and canopy cover **

```{r}
rowMeans(tab, na.rm=TRUE)
terr.pal <- terrain.colors(10)
ggplot(data=raw, aes(x=year, y=o18corr, group=as.factor(Tree_ID))) +
  geom_line(aes(color=as.factor(Tree_ID)))+
  geom_point(aes(color=as.factor(Tree_ID)))+
  scale_color_brewer(palette="Spectral")
```

```{r}
o18_means <- apply(tab, MARGIN = 2, mean, na.rm=TRUE)
c13_means <- apply(tab2, MARGIN=2, mean, na.rm=TRUE)

soildepths <- c(277.5,237.5,591.25,507.5,443.75,372.5,365,335,256.25,356.25)
canopycovers <- c(83.5,76,79.25,83.25,86.25,34,73.25,80.75,81.25,84.25)
```

```{r}
plot(o18_means, soildepths)
```

```{r}
cor.test(o18_means, soildepths)
```

** Compare to GNIP **

```{r}
seas_annual_gnip <- gnip %>% group_by(seas.year) %>% summarise(mean = mean(O18, na.rm = TRUE))
seas_monthly_gnip <- gnip %>% group_by(month) %>% summarise(mean = mean(O18, na.rm = TRUE))
```

```{r}
compare_to_gnip <- tab[19:43,]
```

```{r}
cor.test(seas_annual_gnip[['mean']], compare_to_gnip[['X606']])
```
 606 - p value 0.004875. cor is 0.63
 
 probably irrelevant since only 5 years overlap - 611 p value 0.003535. cor is -0.979
 
 cor is around ~0.3ish, p-value is 0.1 to 0.2ish - 604, 605, 607, 617

```{r}
plot(seas_annual_gnip[['mean']], compare_to_gnip[['X606']])
```

```{r}
DJF_gnip <- subset(gnip, month == 12 | month == 1 | month == 2)
MAM_gnip <- subset(gnip, month == 3 | month == 4 | month == 5)
JJA_gnip <- subset(gnip, month == 6 | month == 7 | month == 8)
SON_gnip <- subset(gnip, month == 9 | month == 10 | month == 11)
```

```{r}
DJF_annual_gnip <- DJF_gnip %>% group_by(seas.year) %>% summarise(mean = mean(O18, na.rm = TRUE))
MAM_annual_gnip <- MAM_gnip %>% group_by(seas.year) %>% summarise(mean = mean(O18, na.rm = TRUE))
JJA_annual_gnip <- JJA_gnip %>% group_by(seas.year) %>% summarise(mean = mean(O18, na.rm = TRUE))
SON_annual_gnip <- SON_gnip %>% group_by(seas.year) %>% summarise(mean = mean(O18, na.rm = TRUE))
de<-data.frame(2002, NA)
names(de)<-c("seas.year","mean")
df<-data.frame(1978, NA)
names(df)<-c("seas.year","mean")
MAM_annual_gnip <- rbind(MAM_annual_gnip, de)
JJA_annual_gnip <- rbind(JJA_annual_gnip, de)
SON_annual_gnip <- rbind(df, SON_annual_gnip)
```

```{r}
cor.test(JJA_annual_gnip[['mean']], compare_to_gnip[['X603']])
```

Notables for DJF: 606, 607, 619
MAM: 606, 612
JJA: 604, 605, 606, 607, 612, 617
SON: 609, negative with 619 very odd

609 is very poor correlation with everything except SON

```{r}
mon_gnip <- subset(gnip, month == 8)
mon_annual_gnip <- mon_gnip %>% group_by(seas.year) %>% summarise(mean = mean(O18, na.rm = TRUE))

de<-data.frame(2002, NA)
names(de)<-c("seas.year","mean")
mon_annual_gnip <- rbind(mon_annual_gnip, de)
```


```{r}
cor.test(mon_annual_gnip[['mean']], compare_to_gnip[['X619']])
```

This SHOULD allow correlations to be for the previous JJA season (double check)
```{r}
compare_to_gnip2 <- tab[20:44,]
```

```{r}
cor.test(JJA_annual_gnip[['mean']], compare_to_gnip2[['X619']])
```


Use GNIP as a treeclim file

```{r}
annual_summary <- raw %>%
  group_by(year) %>%
  summarise(mean = mean(o18corr, na.rm = TRUE))
annual_summary<- as.data.frame((annual_summary))
rownames(annual_summary) <- annual_summary$year
annual_summary[,1] <- NULL
annual_summary$samp.depth <- c(8,8,8,6,8,7,8,7,9,9,9,9,9,7,8,8,8,8,8,8,9,9,9,8,9,9,9,7,9,9,9,8,9,8,9,9,9,9,9,8,9,9,10,10,10,10,10,9,9,9,10,9,10,10,10,10,10,9,9)

gnip_fill <- read.csv("CG_GNIP_filled.csv")
gnip_fill$Date <- parse_date_time(gnip_fill$Date, "mdy")
gnip_fill$month <- month(gnip_fill$Date)
gnip_fill$year <- year(gnip_fill$Date)
gnip_treeclim <- data.frame(gnip_fill$year, gnip_fill$month, gnip_fill$O18)
colnames(gnip_treeclim) <- c("year","month","O18")

treeclim_yrs <- gnip_treeclim %>% group_by(year) %>% summarise(mean = mean(O18))
```

Treeclim will not work with the gnip dataset, as it only takes years with complete climate data. Must figure out a workaround

```{r}
treeclim_months <- gnip_treeclim %>% group_by(month) %>% summarise(mean = mean(O18, na.rm=TRUE))

test_summary <- annual_summary
test_summary$year <- as.numeric(rownames(test_summary))
test_summary$year <- (test_summary$year + 1)
rownames(test_summary) <- test_summary$year
test_summary <- test_summary[,-3]
```


```{r}
dcc(test_summary, gnip_treeclim, .range(-4:12))
```

I THINK:
using rownames(tab) Will correlate tree ring for Oct.1960 - Sep(ish) 1961 to GNIP for 1959-1960
Using rownames(test_summary) will correlate tree ring fro Oct 1960-Sep1961 to GNIP for 1960-61

```{r}
single_tree <- data.frame(mean=tab$X619, samp.depth=1)
rownames(single_tree) = rownames(tab)
```

```{r}
dcc(single_tree, gnip_treeclim, .range(-6:12))
```

**Compare to other factors**

```{r}
ids <- read.ids(master, stc=c(4,2,1))
full_treemeans <- treeMean(master, ids, na.rm=TRUE) 
colMeans(full_treemeans)
```

Tree means for whole cores
```{r}
WidthMeans <- colMeans(full_treemeans, na.rm=TRUE)
```

Tree means for 1960-2018
```{r}
modernWidthMeans <- colMeans(full_treemeans[653:711,], na.rm=TRUE)

widthmeans_df <- data.frame(WidthMeans)
widthmeans.ord <- widthmeans_df[c(4,2,5,9,8,10,6,1,7,3),]

modernWidthMeans_df <- data.frame(modernWidthMeans)
modernwidthmeans.ord <- modernWidthMeans_df[c(4,2,5,9,8,10,6,1,7,3),]
```

```{r}
plot(o18_means, widthmeans.ord)
plot(o18_means, modernwidthmeans.ord)
```

```{r}
cor.test(o18_means, modernwidthmeans.ord)
```

```{r}
plot(c13_means, widthmeans.ord)
plot(c13_means, modernwidthmeans.ord)
```

RW is pretty variable between trees, which could suggest different o18 means between trees. They do not seem correlated, however 

```{r}
DBH <- c(72,75,91,24,86,39,43,65,83,70)
```

```{r}
plot(o18_means, DBH)
plot(c13_means, DBH)
```

```{r}
cor.test(o18_means, DBH)
```

```{r}
o18_sd <- apply(tab, MARGIN = 2, sd, na.rm=TRUE)
c13_sd <- apply(tab2, MARGIN=2, sd, na.rm=TRUE)
```

```{r}
plot(o18_sd, soildepths)
plot(o18_sd, canopycovers)
plot(o18_sd, DBH)
plot(o18_means, o18_sd)
```

```{r}
cor.test(o18_means, o18_sd)
```

```{r}
plot(c13_sd, soildepths)
plot(c13_sd, canopycovers)
plot(c13_sd, DBH)
plot(c13_means, c13_sd)
```

```{r}
o18_annual_sd <- apply(tab, MARGIN = 1, sd, na.rm=TRUE)
c13_annual_sd <- apply(tab2, MARGIN = 1, sd, na.rm=TRUE)
```

```{r}
plot(c(1960:2018), o18_annual_sd)
plot(c(1960:2018), c13_annual_sd)
```

Divide DBH by avg ring width = rough age (smaller number = younger, larger = older)

```{r}
rough_age <- DBH/widthmeans.ord
```

```{r}
plot(o18_means, rough_age)
plot(c13_means, rough_age)
plot(o18_sd, rough_age)
plot(c13_sd, rough_age)
```

```{r}
cor.test(o18_sd, modernwidthmeans.ord)
```

**Prysm**

```{r}
compare_prysm <- annual_summary[1:54,]
prysm_output <- prysm[3:56,]
```

```{r}
cor.test(compare_prysm$mean, prysm_output$prysm)
```

```{r}
newprysm <- read.csv("prysm_SHseason_output.csv")
```

```{r}
compare_prysm <- annual_summary[1:53,]
newprysm_output <- newprysm[4:56,]
```

```{r}
cor.test(compare_prysm$mean, newprysm_output$prysm)
```

```{r}
compare_prysm$year <- rownames(compare_prysm)
prysm_combine <- merge(compare_prysm, newprysm_output, by="year")
prysm_combine$year <- as.numeric(prysm_combine$year)
```

Make line plot for prysm

```{r}
ggplot(prysm_combine, aes(x=year)) + 
  geom_line(aes(y = mean), color = "darkred") + 
  geom_line(aes(y = prysm), color="steelblue", linetype="twodash") +
  theme(legend.position="bottom")
```

```{r}
prysm_full <- read.csv("prysm_full.csv")
prysm_full$date <- parse_date_time(prysm_full$date, "mdy")
prysm_full <- prysm_full[25:672,]
```

```{r}
annual_mean_o18 <- annual_summary[1:54,]
```

Compare PRYSM monthly output:

```{r}
jan_prysm <- subset(prysm_full, month == 1)
feb_prysm <- subset(prysm_full, month == 2)
mar_prysm <- subset(prysm_full, month == 3)
apr_prysm <- subset(prysm_full, month == 4)
may_prysm <- subset(prysm_full, month == 5)
jun_prysm <- subset(prysm_full, month == 6)
jul_prysm <- subset(prysm_full, month == 7)
aug_prysm <- subset(prysm_full, month == 8)
sep_prysm <- subset(prysm_full, month == 9)
oct_prysm <- subset(prysm_full, month == 10)
nov_prysm <- subset(prysm_full, month == 11)
dec_prysm <- subset(prysm_full, month == 12)
```

```{r}
cor.test(annual_mean_o18$mean, dec_prysm$d18o)
```

```{r}
annual_mean_o18$year <- rownames(annual_mean_o18)
```

```{r}
temp_prysm <- merge(annual_mean_o18, jan_prysm, by="year")
temp_prysm$year <- as.numeric(temp_prysm$year)
ggplot(temp_prysm, aes(x=year)) + 
  geom_line(aes(y = mean), color = "darkred") + 
  geom_line(aes(y = d18o), color="steelblue", linetype="twodash") 
```

Compare growing season prysm output to data:

```{r}
growingseason_prysm <- subset(prysm_full, month == 12 | month ==1 | month == 2)
```

```{r}
gs_summary <- growingseason_prysm %>% group_by(seasYear) %>% summarise(d18o = mean(d18o, na.rm = TRUE))
gs_summary <- gs_summary[2:55,]
colnames(gs_summary) <- c("year","d18o")
```

```{r}
cor.test(annual_mean_o18$mean, gs_summary$d18o)
```


```{r}
temp_prysm_gs <- merge(annual_mean_o18, gs_summary, by="year")
temp_prysm_gs$year <- as.numeric(temp_prysm_gs$year)
ggplot(temp_prysm_gs, aes(x=year)) + 
  geom_line(aes(y = mean), color = "darkred") + 
  geom_line(aes(y = d18o), color="steelblue", linetype="twodash") 
```

Compare individual tree series to PRYSM:

```{r}
tab_prysm <- tab[1:54,]
```

```{r}
cor.test(tab_prysm$X619, feb_prysm$d18o)
```

Year(?) - 617

feb - 605
Mar-604 (negative), 606(negative), 607(negative), 617 (negative)
Apr - 604(negative), 609
May - 609
Jun - 604(negative), 605(negative), 611
Jul - 611
Aug - 605, 609
sep - 604(negative), 617(negative)

Plotting:

```{r}
tab_prysm$year <- as.numeric(rownames(tab_prysm))
tab_prysm_comb <- merge(tab_prysm[1:53,], newprysm_output, by="year") 
```

```{r}
ggplot(tab_prysm_comb, aes(x=year)) + 
  geom_line(aes(y = X605), color = "darkred") + 
  geom_line(aes(y = prysm), color="steelblue", linetype="twodash")
```

Tasks as of 7/5:

1. Find flyers
Identified years to remove: 619 1973-1978; 609 1982; 606 1997
Others that could possibly be removed

2. See how similar RW is between trees

It's variable, but that variability is not necessarily correlated with 18O

3. Run oxygen thru dplR in cofecha


Figure out what to do for NAs in oxygen series.
Then:
Detrend
corr.rwl.seg
chron
corr.series.seg(?)
rwl.stats
rwl.stats.running

```{r}
filled <- read.csv("filled_18o.csv")
rownames(filled) <- filled[,1]
filled[,1] <- NULL
detrended_18o <- detrend(filled, method="Spline")
```

```{r}
chron(detrended_18o)
```

```{r}
rwl.stats(filled)
```

```{r}
rwl.stats(detrended_18o)
```

```{r}
crn.plot(chron(detrended_18o))
```


4. Combine groupings -  create chronologies - analyze

Reminder of groups:

603 - 619, 612, 605
619 - 603, 605, 612
612 - 619, 603

605 - 619, 604, 603 ---------------------Link between groups maybe?

604 - 607, 609, 605
607 - 604, 611
609 - 604
611 - 607

606 - 617**
617 - 606**



```{r}
downslope_sites <- subset(tab, select=c("X603","X605","X612","X619"))
upslope_sites <- subset(tab, select=c("X604","X607","X609","X611"))
```

```{r}
downslope_raw <- raw %>% filter(Tree_ID == 603 | Tree_ID == 605 |Tree_ID == 612 |Tree_ID == 619)
upslope_raw <- raw %>% filter(Tree_ID == 604 | Tree_ID == 607 |Tree_ID == 609 |Tree_ID == 611)
```

```{r}
ggplot(data=downslope_raw, aes(x=year, y=o18corr, group=as.factor(Tree_ID))) +
  geom_line(aes(color=as.factor(Tree_ID)))+
  geom_point(aes(color=as.factor(Tree_ID)))+
  scale_color_brewer(palette="Spectral")
```

```{r}
ggplot(data=upslope_raw, aes(x=year, y=o18corr, group=as.factor(Tree_ID))) +
  geom_line(aes(color=as.factor(Tree_ID)))+
  geom_point(aes(color=as.factor(Tree_ID)))+
  scale_color_brewer(palette="Spectral")
```

```{r}
upslope_summary <- upslope_raw %>% group_by(year) %>% summarise(mean = mean(o18corr, na.rm = TRUE))
downslope_summary <- downslope_raw %>% group_by(year) %>% summarise(mean = mean(o18corr, na.rm = TRUE))
upslope_summary<- as.data.frame((upslope_summary))
rownames(upslope_summary) <- upslope_summary$year
upslope_summary[,1] <- NULL
upslope_summary$samp.depth <- c(4,3,4,4,4,4,4,3,4,4,4,4,4,4,4,3,3,3,4,4,4,4,2,3,3,3,3,2,3,3,3,3,3,3,3,3,3,3,3,3,3,4,4,4,4,4,4,3,3,3,4,4,4,4,4,4,4,4,4)
downslope_summary<- as.data.frame((downslope_summary))
rownames(downslope_summary) <- downslope_summary$year
downslope_summary[,1] <- NULL
downslope_summary$samp.depth <- c(3,4,3,2,3,2,3,3,4,4,4,4,4,1,2,3,3,3,2,3,4,3,4,3,4,4,4,4,4,4,4,3,4,3,4,4,4,4,4,4,4,3,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,3,3)
```



Convert to correct format, add sample depth.

```{r}
dcc(upslope_summary, silo, .range(-1:12))
```

```{r}
dcc(downslope_summary, silo, .range(-1:12))
```


Rework gnip tool for this

```{r}
newsilo <- silo[37:744,]
```
This needs Edit:
```{r}
climcorrs <- function(trees){
corrs <- rep(NA, 10)
for (m in 1:12){
mon_silo <- subset(newsilo, month == m)
test <- cor.test(mon_annual_silo[['mean']], trees)
corrs[m] <- test$estimate
}
return(corrs)
}
```


```{r}
gnipcorrs2 <- function(trees){
corrs <- rep(NA, 10)
for (m in 1:12){
mon_gnip <- subset(gnip_outlier, month == m)
test <- cor.test(mon_gnip[['O18']], trees)
corrs[m] <- test$estimate
}
return(corrs)
}
```

```{r}
gnip_group_corrs <- data.frame(row.names = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
```

```{r}
slope_compare <- downslope_summary[20:43,]
next_compare <- downslope_summary[19:42,]
gnip_group_corrs$down_prev <- gnipcorrs2(slope_compare$mean)
gnip_group_corrs$down_next <- gnipcorrs2(next_compare$mean)
```

Plot the gnip vs. isotopes:
```{r}
mon_gnip <- subset(gnip_outlier, month == 2)
plot(next_compare$mean, mon_gnip$O18)
```

Standardize groups before averaging

```{r}
standardize_up <- function(num) {
  up_std <- upslope_sites[[num]] - up_means[[num]]
  return(up_std)
}
standardize_down <- function(num) {
  down_std <- downslope_sites[[num]] - down_means[[num]]
  return(down_std)
}
```

```{r}
up_means <- colMeans(upslope_sites, na.rm=TRUE)
down_means <- colMeans(downslope_sites, na.rm=TRUE)
```

```{r}
stand_up_o18 <- data.frame(row.names = rownames(tab))
for (x in c("X604","X607","X609","X611")){
  stand_up_o18[[x]] <- standardize_up(x)
}
```

```{r}
stand_down_o18 <- data.frame(row.names = rownames(tab))
for (x in c("X603","X605","X612","X619")){
  stand_down_o18[[x]] <- standardize_down(x)
}
```


```{r}
downslope_std_summary <- data.frame(rowMeans(stand_down_o18, na.rm=TRUE))
upslope_std_summary <- data.frame(rowMeans(stand_up_o18, na.rm=TRUE))
colnames(upslope_std_summary) <- "mean"
colnames(downslope_std_summary) <- "mean"
```

```{r}
plot(rownames(upslope_std_summary), upslope_std_summary$mean, type = "b", frame = FALSE, pch = 19, 
     col = "red", xlab = "x", ylab = "y")
# Add a second line
lines(rownames(upslope_std_summary), downslope_std_summary$mean, pch = 18, col = "blue", type = "b", lty = 2)
```


Correlations will probably be the same but it's worth doing them again with these series

```{r}
slope_compare <- downslope_std_summary[19:42,]
```

```{r}
gnipcorrs2(slope_compare)
```

climcorrs


5. Standardize 18O

```{r}
standardize_o18 <- function(num) {
  new_std <- tab[[num]] - o18_means[[num]]
  return(new_std)
}
```

```{r}
stand_o18 <- data.frame(row.names = rownames(tab))
for (x in c("X603","X604","X605","X606","X607","X609","X611","X612","X617","X619")){
  stand_o18[[x]] <- standardize_o18(x)
}
```

```{r}
pheatmap(as.matrix(stand_o18), scale="none", cluster_rows = FALSE, cluster_cols = FALSE)
```

```{r}
stand_o18$date <- as.integer(rownames(stand_o18))
ggplot(stand_o18, aes(x=date)) + 
  geom_line(aes(y = X603), color = "darkred") + 
  geom_line(aes(y = X604), color="steelblue") +
  geom_line(aes(y = X605), color = "orange") + 
  geom_line(aes(y = X606), color="aquamarine") +
  geom_line(aes(y = X607), color = "brown") + 
  geom_line(aes(y = X609), color="deepskyblue")+
  geom_line(aes(y = X611), color = "deeppink") + 
  geom_line(aes(y = X612), color="darkviolet")+
  geom_line(aes(y = X617), color = "coral") + 
  geom_line(aes(y = X619), color="chocolate")
```

6. GNIP pearson correlations

```{r}
compare_to_gnip <- tab[19:42,]
```

```{r}
gnipcorrs <- function(tree){
corrs <- rep(NA, 10)
for (m in 1:12){
mon_gnip <- subset(gnip, month == m)
mon_annual_gnip <- mon_gnip %>% group_by(year) %>% summarise(mean = mean(O18, na.rm = TRUE))
test <- cor.test(mon_annual_gnip[['mean']], compare_to_gnip[[tree]])
corrs[m] <- test$estimate
}
return(corrs)
}
```

```{r}
yr.to.yr.corrs <- data.frame(row.names = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
nextyr.corrs <- data.frame(row.names = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
```


```{r}
nextyr.corrs$X619 <- gnipcorrs("X619")
```


7. Compare PRYSM to CG GNIP
-Possibly use isoGSM for PRYSM


```{r}
prysm_full$Date <- prysm_full$date
prysm_vs_gnip <- merge(prysm_full, gnip, by="Date")
```

```{r}
ggplot(prysm_vs_gnip, aes(x=Date)) + 
  geom_line(aes(y = O18), color = "darkred") + 
  geom_line(aes(y = d18o), color="steelblue", linetype="twodash")
```

```{r}
cor.test(prysm_vs_gnip$d18o, prysm_vs_gnip$O18)
```

Cluster analysis

```{r}
newfill <- read.csv("filled_for_cluster_18o.csv")
newfill <- scale(newfill)
clusterset <- data.frame(t(newfill))
clusterset <- clusterset[2:10,]
```

```{r}
kmeans(clusterset, centers=4, nstart=10)
```

4,6,7,17
3,5,9,12,19

Try these new groupings (figure out where to put 11)

```{r}
group1_sites <- subset(tab, select=c("X603","X605","X609","X612","X619"))
group2_sites <- subset(tab, select=c("X604","X606","X607","X617"))
group1_raw <- raw %>% filter(Tree_ID == 603 | Tree_ID == 605 | Tree_ID == 609 |Tree_ID == 612 |Tree_ID == 619)
group2_raw <- raw %>% filter(Tree_ID == 604 | Tree_ID == 606 |Tree_ID == 607 |Tree_ID == 617)
```

```{r}
ggplot(data=group1_raw, aes(x=year, y=o18corr, group=as.factor(Tree_ID))) +
  geom_line(aes(color=as.factor(Tree_ID)))+
  geom_point(aes(color=as.factor(Tree_ID)))+
  scale_color_brewer(palette="Spectral")
```

```{r}
ggplot(data=group2_raw, aes(x=year, y=o18corr, group=as.factor(Tree_ID))) +
  geom_line(aes(color=as.factor(Tree_ID)))+
  geom_point(aes(color=as.factor(Tree_ID)))+
  scale_color_brewer(palette="Spectral")
```
Note: samp depths are wrong, will not work with dcc
```{r}
group2_summary <- group2_raw %>% group_by(year) %>% summarise(mean = mean(o18corr, na.rm = TRUE))
group1_summary <- group1_raw %>% group_by(year) %>% summarise(mean = mean(o18corr, na.rm = TRUE))
group2_summary<- as.data.frame((group2_summary))
rownames(group2_summary) <- group2_summary$year
group2_summary[,1] <- NULL
group2_summary$samp.depth <- c(4,3,4,4,4,4,4,3,4,4,4,4,4,4,4,3,3,3,4,4,4,4,2,3,3,3,3,2,3,3,3,3,3,3,3,3,3,3,3,3,3,4,4,4,4,4,4,3,3,3,4,4,4,4,4,4,4,4,4)
group1_summary<- as.data.frame((group1_summary))
rownames(group1_summary) <- group1_summary$year
group1_summary[,1] <- NULL
group1_summary$samp.depth <- c(3,4,3,2,3,2,3,3,4,4,4,4,4,1,2,3,3,3,2,3,4,3,4,3,4,4,4,4,4,4,4,3,4,3,4,4,4,4,4,4,4,3,4,4,4,4,4,4,4,4,4,4,4,4,4,4,4,3,3)
```

Plot groups against each other

```{r}
plot(rownames(group1_summary),group1_summary$mean,type="l",col="red")
lines(rownames(group1_summary),group2_summary$mean,col="green")
```


corrs w/gnip

```{r}
new_group_corrs <- data.frame(row.names = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
```

```{r}
g1slope_compare <- group1_summary[20:43,]
g1next_compare <- group1_summary[19:42,]
g2slope_compare <- group2_summary[20:43,]
g2next_compare <- group2_summary[19:42,]
new_group_corrs$g1_prev <- gnipcorrs2(g1slope_compare$mean)
new_group_corrs$g1_next <- gnipcorrs2(g1next_compare$mean)
new_group_corrs$g2_prev <- gnipcorrs2(g2slope_compare$mean)
new_group_corrs$g2_next <- gnipcorrs2(g2next_compare$mean)
```

Plot the gnip vs. isotopes:
```{r}
mon_gnip <- subset(gnip_outlier, month == 11)
gnip_group_plotting <- data.frame(mon_gnip$O18, mon_gnip$year, g2next_compare$mean)
colnames(gnip_group_plotting) <- c("O18","year","mean")
ggplot(data=gnip_group_plotting, aes(x=O18, y=mean, color=year))+
geom_point()+
geom_text(aes(label=year),hjust=0.5, vjust=-0.5)
```









Carbon Suess effect correction





















** Don't even touch this yet **
May need to adjust dates due to SH dating (2018 is Dec 2018 - May 2019)

```{r}
silo <- read.csv("silo_monthly_1958_2019.csv")
```

```{r}
annual_summary <- raw %>%
  group_by(year) %>%
  summarise(mean = mean(o18corr, na.rm = TRUE))
annual_summary<- as.data.frame((annual_summary))
rownames(annual_summary) <- annual_summary$year
annual_summary[,1] <- NULL
annual_summary$samp.depth <- c(8,8,8,6,8,7,8,7,9,9,9,9,9,7,8,8,8,8,8,8,9,9,9,8,9,9,9,7,9,9,9,8,9,8,9,9,9,9,9,8,9,9,10,10,10,10,10,9,9,9,10,9,10,10,10,10,10,9,9)
```

```{r}
dcc(annual_summary, silo, .range(-1:12))
```

```{r}
seascorr(annual_summary, silo)
```

```{r}
plot(seascorr(annual_summary, silo))
```


```{r}
newdf <- as.data.frame(tab$X619)
rownames(newdf) <- rownames(tab)
newdf$samp.depth <- 1
dcc(newdf, silo, .range(-1:12))
```











```{r}
ERA_precip <- brick("ERA5_precip.gri")
ERA_t2m <- brick("ERA5_t2m.gri")
```

Create yearly averages

```{r}
datC <- brick("ERA5_t2m.gri")
```

```{r}
yr <- "1955" # Year you want to subset

sel_y <- names(datC)[substr(names(datC), 2, 5) %in% yr] #Creates true/false field of rasters for that yr
#use that index to subset the raster brick
datY <- subset(datC, sel_y, value=T) #subsets the dataset of only those that are true
tail(names(datY), 30)
mean <- calc(datY, fun = mean)
```

```{r}
annual_ERA_t2m <- raster(ncol=720, nrow=360, xmn=0.125, xmx=180.125, ymn=-89.875, ymx=0.125, crs=CRS("+proj=longlat +datum=WGS84"))
for (yr in 1960:2018){
  sel_y <- names(datC)[substr(names(datC), 2, 5) %in% yr] #Creates true/false field of rasters for that yr
#use that index to subset the raster brick
datY <- subset(datC, sel_y, value=T) #subsets the dataset of only those that are true
mean <- calc(datY, fun = mean)
annual_ERA_t2m <- brick(annual_ERA_t2m, mean)
}
```



Combine all of the ERA data together into a single stack (AND CROP TO 1958 - 2013! Change this if needed)
```{r}
ERA_t2m_sub <- subset(ERA_t2m, 109:828)
ERA_precip_sub <- subset(ERA_precip, 109:828)
```


```{r}
ar.t2m <- as.array(ERA_t2m_sub)
ar.prcp <- as.array(ERA_precip_sub)
```




```{r}
apply(ar.t2m, 1:2, rcorr, MRD_18o, type="pearson")
```

This is in list form basically. The cells go like this

1   5   9   13  17
2   6   10  14  18
3   7   11  15  19
4   8   12  16  20

```{r}
z <- apply(ar.prcp, 1:2, rcorr, MRD_18o, type="pearson")
```

correlation = z[[x]][[1]][2]
significance = z[[x]][[3]][2]

Up next: create corr and sig arrays

Change (67,89) to new extent:

```{r}
corr = array(0, dim=c(360,720) )
sig = array(0, dim=c(360,720) )
```

Change 5963 to number of cells in new extent (ensure routing is correct):

Must transpose corr and sig to get right dimensions
```{r}
for (x in 1:259200)
{
corr[x] <- z[[x]][[1]][2]
sig[x] <- z[[x]][[3]][2]
}
corr <- t(corr)
sig <- t(sig)
```




Make it the right structure(?)

```{r}
l <- list(corr, sig)
a <- do.call(rbind, l)
library(abind)
a <- abind(l, along=3)
a
```

Change this brick call to get correct resolution, extent, etc
```{r}
r <- brick(a, xmn=0.125, xmx=180.125, ymn=-89.875, ymx=0.125, crs=CRS("+proj=longlat +datum=WGS84"))
r
```
0.125, 180.125, -89.875, 0.125  (xmin, xmax, ymin, ymax)
```{r}
plot(r)
```

Create raster of corrs:

```{r}
writeRaster(r, "o18_prcp_corr")
```





