---
title: "MRD_workspace_2"
author: "Zack Grzywacz"
date: "July 15, 2021"
output: html_document
---


```{r}
library(lubridate)
library(lattice)
library(rnaturalearth)
library(rnaturalearthdata)
library(rasterVis)
library(seas)
library(s2dverification)
library(reshape2)
library(dplyr)
library(Hmisc)
library(ggplot2)
library(viridis)
library(ggExtra)
library(grid)
library(pheatmap)
library(treeclim)
library(raster)
library(dplR)
library(scPDSI)
```


```{r}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
```

```{r}
raw <- read.csv("raw_isotopes.csv") # Raw isotope data
raw$year <- substrRight(raw$sample_ID, 4)
raw$year <- strtoi(raw$year)

annual_summary <- raw %>% group_by(year) %>% summarise(mean = mean(o18corr, na.rm = TRUE)) # Combined Isotope data

tab <- read.csv("tabulated_18o.csv") # Oxygen data tabulated
rownames(tab) <- tab[,1]
tab[,1] <- NULL

tab2 <- read.csv("tabulated_13c.csv") # carbon data tabulated
rownames(tab2) <- tab2[,1]
tab2[,1] <- NULL

gnip <- read.csv("CG_GNIP.csv")
gnip$Date <- parse_date_time(gnip$Date, "mdy")
gnip$month <- month(gnip$Date)
gnip$year <- year(gnip$Date)
for (i in c(1:288)){
if (gnip$month[i] < 9){
  gnip$seas.year[i] = (gnip$year[i] - 1)
}
else{
  gnip$seas.year[i] = gnip$year[i]
}
}

gnip_outlier <- gnip
gnip_outlier$O18[gnip_outlier$O18 > 1] <- NA

vpd <- read.csv("vpd_monthly_1959_2019.csv")
vpd <- vpd[13:720,]

clim <- read.csv("silo_monthly_1959_2019.csv")
```

Standardize 18O series

```{r}
std_18o <- scale(tab)
```

Corr matrix

```{r}
cor(std_18o, use = "complete.obs")
```

The same - duh.

```{r}
std_summary <- rowMeans(std_18o, na.rm=TRUE)
```


Groupings

```{r}
downslope_sites <- subset(std_18o, select=c("X603","X605","X612","X619"))
upslope_sites <- subset(std_18o, select=c("X604","X607","X609","X611"))
cluster_group1 <- subset(std_18o, select=c("X603","X605","X609","X612","X619"))
cluster_group2 <- subset(std_18o, select=c("X604","X606","X607","X617"))
downslope_summary <- data.frame(rowMeans(downslope_sites, na.rm=TRUE))
colnames(downslope_summary) <- "mean"
upslope_summary <- data.frame(rowMeans(upslope_sites, na.rm=TRUE))
colnames(upslope_summary) <- "mean"
group1_summary <- data.frame(rowMeans(cluster_group1, na.rm=TRUE))
colnames(group1_summary) <- "mean"
group2_summary <- data.frame(rowMeans(cluster_group2, na.rm=TRUE))
colnames(group2_summary) <- "mean"
```



Make correlation function that corrs parameter plus lag
Take all factors -> output plot/table


Change this slightly to allow for nextyr corrs as well
```{r}
o18_corrs <- function(df, trees){
  newclim <- df[13:720,]
  nextclim <- df[25:732,]
  output <- data.frame(Clim="", Month="", Corr="", p="")
  for (param in c(3:ncol(df))){
    tempdf <- data.frame(Clim=colnames(df[param]), Month=c(6:12), Corr="", p="")
    tempdf2 <- data.frame(Clim=colnames(df[param]), Month=c(1:5), Corr="", p="")
    for (m in 6:12){
      n = (m-5)
      mon_sub <- subset(newclim, month == m)
      test <- cor.test(mon_sub[[param]], trees)
      tempdf$Corr[n] <- test$estimate
      tempdf$p[n] <- test$p.value
    }
    for (m in 1:5){
      mon_sub <- subset(nextclim, month == m)
      test <- cor.test(mon_sub[[param]], trees)
      tempdf2$Corr[m] <- test$estimate
      tempdf2$p[m] <- test$p.value
    }
  output <- rbind(output, tempdf, tempdf2)
  }
  output <- output[-1,]
  output$Corr <- as.numeric(output$Corr)
  output$p <- as.numeric(output$p)
  output$Month <- c("Jun","Jul","Aug","Sep","Oct","Nov","Dec","Jan","Feb","Mar","Apr","May")
  return(output)
}
```


```{r}
o18_corrs(clim, annual_summary$mean)
```



Figure out scPDSI

Take this output and add it to clim data frame

```{r}
pdsi_output <- pdsi(clim$precip, clim$pet, start=1959, sc=TRUE)
```

```{r}
clim$pdsi <- pdsi_output$X
clim$phdi <- pdsi_output$PHDI
clim$weight.pdsi <- pdsi_output$WPLM
```

```{r}
o18_corrs(clim, annual_summary$mean)
```


```{r}
newdf <- o18_corrs(clim, std_summary)
newdf$Month <- factor(newdf$Month, levels=unique(newdf$Month))
gg <- ggplot(newdf, aes(x=Month, y=Clim, fill=Corr)) + geom_tile(color="white", size=0.1) + scale_fill_viridis(name="# Events") + coord_equal() + labs(x=NULL, y=NULL, title="Monthly Correlations")+geom_point(data=newdf[newdf$p<0.05,])
gg
```


Groupings:

```{r}
newdf <- o18_corrs(clim, downslope_summary$mean)
newdf$Month <- factor(newdf$Month, levels=unique(newdf$Month))
gg <- ggplot(newdf, aes(x=Month, y=Clim, fill=Corr)) + geom_tile(color="white", size=0.1) + scale_fill_viridis(name="# Events") + coord_equal() + labs(x=NULL, y=NULL, title="Downslope Monthly Correlations")+geom_point(data=newdf[newdf$p<0.05,])
gg
```


```{r}
newdf <- o18_corrs(clim, upslope_summary$mean)
newdf$Month <- factor(newdf$Month, levels=unique(newdf$Month))
gg <- ggplot(newdf, aes(x=Month, y=Clim, fill=Corr)) + geom_tile(color="white", size=0.1) + scale_fill_viridis(name="# Events") + coord_equal() + labs(x=NULL, y=NULL, title="Upslope Monthly Correlations")+geom_point(data=newdf[newdf$p<0.05,])
gg
```


Individual Trees:

```{r}
newdf <- o18_corrs(clim, std_18o[,10])
newdf$Month <- factor(newdf$Month, levels=unique(newdf$Month))
gg <- ggplot(newdf, aes(x=Month, y=Clim, fill=Corr)) + geom_tile(color="white", size=0.1) + scale_fill_viridis(name="# Events") + coord_equal() + labs(x=NULL, y=NULL, title="619")+geom_point(data=newdf[newdf$p<0.05,])
gg
```

```{r}
std_18o
```


C Seuss correction
Ask Amy - seems complicated

Corrected 13C = (tree ring 13C) - (atmosphere d13CO2 + pre-industrial d13CO2)
Pre-industrial is ~ -6.5 (I think)

Load in atmosphere data -> add pre-industrial -> subtract from tree ring 

```{r}
testdf<- read.csv("monthly_flask_c13_mlo.csv")
testdf <- testdf[c(1:2,9)]
c13_atm <- testdf[9:512,]
colnames(c13_atm) <- c("Yr","Mn","C13")
c13_atm$C13 <- as.numeric(c13_atm$C13)
c13_atm[1,3] <- NA
```

```{r}
annual_c13_atm <- c13_atm %>% group_by(Yr) %>% summarise(c13 = mean(C13, na.rm=TRUE))
annual_c13_atm$corrected <- (annual_c13_atm$c13 - 6.5)
carbon_correction <- tab2[21:59,]
split_annual_c13 <- annual_c13_atm[1:39,]
new_c13 <- carbon_correction - split_annual_c13$corrected
new_c13
```

```{r}
std_13c <- data.frame(scale(new_c13))
std_13c <- std_13c[,-4]
c13_summary <- rowMeans(std_13c, na.rm=TRUE)
c13_summary
```

```{r}
plot(c13_summary)
```

